---
title: "3_dischargeAnalysis"
output: html_document
date: "2023-03-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
library(measurements)
library(sf)
library(lubridate)
library(grDevices)
library(mapview)
library(extrafont)
library(ggpubr)
library(ggmap)
library(RgoogleMaps)
library(broom)
library(feather)
library(tidyhydat)
library(sp)
library(data.table)
library(ggalluvial)
library(patchwork)
library(magick)
library(units)
library(Kendall)
library(ggspatial)
library(dtplyr)
#Import libraries for Random Forest
library(caret) 
library(e1071)
library(Boruta)
library(tidymodels)
library(skimr)
library(vip)
```

# Set filenames and directories and constants
```{r}
# dates for version control
todayDate  = "20230324" # the first data join phase

# Names of files and folders for reflectance data
import.filePath = "C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/Data/GEE Downloads"


# intermediate working directory
int.wd="C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/Data/intermediaryDownloads"


#Name of file and folder for lake shapefiles & island polygon shapefiles
shapeFiles.filePath = "C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/Data/shapeFiles"
lakes.shapeFile = "mackenzieGoodLakes.shp"
islands.shapeFile = "vectorIslandArea2.shp"
import.sword = "na_sword_reaches_hb82_v14.shp"
setwd(shapeFiles.filePath)
lakes.sf = st_read(lakes.shapeFile)
islands.sf=st_read(islands.shapeFile)

images.wd = "C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/images"

# import Marsh & Hey Validation
val.wd = "C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/Data/MarshHey1988"
val.filename = "MarshHey1998_prj.shp"

# River waterlevel data
wsc.wd = "C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/Data/dischargeData"

# ArcticGRO sediment & level data at Arctic Red River 
sed.wd="C:/Users/whyana/OneDrive - University of North Carolina at Chapel Hill/DocumentsLaptop/001_ Graduate School/Research/Connectivity/Mackenzie/Data"
sed.file= "ArcticGROWaterQuality.csv"
```

# Import river centerlines and set the projection for all future plots, import classifications
```{r}
crs.plot = "+proj=tcea +lon_0=-134.3847656 +datum=WGS84 +units=m +no_defs"
setwd(shapeFiles.filePath)
study.area.large=cbind.data.frame(lon=c(-136.80, -136.80, -133.47, -133.47), 
                 lat=c(67.25, 69.55, 69.55, 67.46)) %>% 
  st_as_sf(coords=c("lon", "lat")) %>% st_set_crs(4326) %>% st_bbox() %>% st_as_sfc() %>% 
  st_transform(crs = crs.plot)
mack.basin.large = st_read(import.sword) %>% 
  st_transform(crs = crs.plot) %>% 
  st_intersection(study.area.large) %>% dplyr::filter(width>90)

# import classifications
setwd(int.wd)
all.classified.filter = read_feather(paste0("final.class_", todayDate, ".feather")) 

```

# Import river water level data
```{r}
setwd(wsc.wd)
level.files = list.files(pattern="*.csv")

readfun <- function(x) {
  dataset <- fread(x,header=TRUE, sep=",", skip=1)
  #setnames(dataset,c("Name1","Name2"))
  return(dataset)
}

# Import water level data keeping good years between 1984-present, correct water levels to CGG05 projection (Véronneau, 2006) - values in Table 2.3 https://central.bac-lac.gc.ca/.item?id=TC-AEU-30267&op=pdf&app=Library&oclc_number=802293902
level_data <- rbindlist(lapply(level.files,readfun)) %>% 
  filter(PARAM==2) %>% #keep only water level, not discharge
  mutate(month=month(Date),
         year=year(Date),
         doy = yday(Date)) %>% 
  filter(month>=4 & month <=9 & year>=1984) %>% 
  filter(!(ID=="10LC012" & year>=2012 & year<=2013)) %>% 
  filter(!(ID=="10MC023" & year <=1999)) %>% 
  mutate(Value = case_when(
    ID == "10LC002" ~ Value-10.856,
    ID == "10LC012" ~ Value-9.822,
    ID == "10LC013" ~ Value-9.713,
    ID == "10LC014" ~ Value-0.024,
    ID == "10LC021" ~ Value-9.056,
    ID == "10MC002" ~ Value+0.074,
    ID == "10MC003" ~ Value-10.056,
    ID == "10MC008" ~ Value-10.346,
    ID == "10MC023" ~ Value-10.603,
    ID == "10MC011" ~ Value-9.213
  ))

station.ids = unique(level_data$ID)

level.locations = hy_stations(station=station.ids) %>% 
  st_as_sf(coords=c("LONGITUDE","LATITUDE")) %>% st_set_crs(4326) %>% 
  st_transform(crs = crs.plot) %>% 
  rename(ID=STATION_NUMBER)
```

#  Complete Functional Sill Elevation Calculation 
```{r}
# Calculate mean water level at each WSC station
level.prep = level_data %>% 
   select(ID, Date, Value) %>% 
   spread(ID, Value) %>% 
    na.omit() %>% select(-Date) %>% colMeans() 
# Get station numbers for all the stations
col.names = level_data %>% 
   select(ID, Date, Value) %>% 
   spread(ID, Value) %>% 
   na.omit() %>% select(-Date) %>% colnames()

# join the water level to the station location information
level.prep2 = cbind.data.frame(level.prep %>% as_tibble(), col.names) %>% 
  rename(ID = col.names) %>% 
  left_join(level.locations %>% 
              select(STATION_NAME, ID, geometry), 
            by="ID") %>% 
  st_as_sf()

# calculate a distance matrix between all stations
dist.matrix = st_distance(level.prep2)
dimnames(dist.matrix) = list(col.names, col.names)
dist.df = t(combn(colnames(dist.matrix), 2))
dist.df = data.frame(dist.df, dist = dist.matrix[dist.df])

# calculate a  difference matrix between mean water levels at each station
dif.matrix = dist(level.prep2$value, diag=T, upper=T) %>% as.matrix()
dimnames(dif.matrix) = list(col.names, col.names)
dif.df = t(combn(colnames(dif.matrix), 2))
dif.df = data.frame(dif.df, dist = dif.matrix[dif.df]) %>% rename(dif = dist)

# combine distance and difference matrices
dist.dif.df = dist.df %>% left_join(dif.df, by=c("X1", "X2")) %>% as_tibble()

# plot the relationship between distance between stations and differences in water level
dist.dif.df %>% 
  filter(X1 != "10MC002" &  X1 != "10LC014" & X1!="10LC021" & X1 != "10MC008") %>% 
  filter(X2 != "10MC002" &  X2 != "10LC014"& X2!="10LC021" & X2 != "10MC008") %>% 
  mutate(dist = as.numeric(conv_unit(dist,from="m", to="km"))) %>% 
  ggplot()+geom_point(aes(x=dist, y=dif))+theme_classic()+
  theme(axis.text = element_text(size=12),
        axis.title = element_text(size=12, face="bold"))+
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1.0, 1.25, 1.5))+
  geom_smooth(aes(x=dist, y= dif), method="lm", se=F)+
  xlab("distance between station pairs [km]")+ylab("difference in average\nwater level between station pairs [m]")+labs(color="WSC Station\nNumber")
setwd(images.wd)
ggsave(paste0(todayDate,"_levelDiff.png"), width = 6, height = 5, units = "in")

# filter out the stations that are upstream of the delta, since they behave differently
dist.dif.df.filt = dist.dif.df %>% 
  filter(X1 != "10MC002" &  X1 != "10LC014" & X1!="10LC021" & X1 != "10MC008") %>% 
  filter(X2 != "10MC002" &  X2 != "10LC014"& X2!="10LC021" & X2 != "10MC008") 
# Calculate the relationship between distance and difference in water level
dist.dif.mod = lm(dif ~ as.numeric(dist), data =dist.dif.df.filt )
mod.int = dist.dif.mod$coefficients[[1]]
mod.slope = dist.dif.mod$coefficients[[2]]
mod.slope
# For each station select lakes that are within 50km
station.buffer = level.locations %>% st_buffer(50000) %>% # buffer by 50km
  select(ID, STATION_NAME, geometry) %>% 
  filter(ID != "10MC002" &  ID != "10LC014" &
           ID != "10LC021" & ID != "10MC008") 

lakes.proj = lakes.sf %>% st_transform(crs.plot) %>% 
  select(OBJECTID, geometry)
lake.list = st_intersects(lakes.proj, station.buffer)
lakes.combo = lakes.proj[lengths(lake.list)>0,] %>% st_join(station.buffer) 


# get distances between each lake and the relevant station
lakes.point = lakes.combo %>% st_centroid()
station.point = lakes.combo %>% as_tibble() %>% select(ID) %>% 
  left_join(level.locations, by="ID") %>% 
  st_as_sf() %>% select(ID)

lake.station.dist = st_distance(lakes.point, station.point, by_element=T) %>% as_tibble()

lakes.station.dist = cbind.data.frame(lakes.combo, lake.station.dist) %>% 
  as_tibble() %>% rename(dist.m = value) %>% 
  mutate(dist_error = dist.m * mod.slope)

# combine the lake classifications with the discharge data - slow, need to speed up
nest.sill = all.classified.filter %>% 
  left_join(lakes.station.dist %>% as_tibble() %>% select(-geometry), by="OBJECTID") %>% 
  filter(!is.na(ID)) %>% 
  mutate(.pred_class = as.numeric(as.character(.pred_class))) %>% 
  filter(.pred_class !=1) %>% # remove the middle 'catch-all' class
  left_join(level_data %>% rename(date=Date, Value.0 = Value) %>% 
              select(ID, Value.0, date) %>% mutate(date=as_date(date)), by=c("date", "ID"))%>% 
  na.omit() %>% group_by(OBJECTID, ID, 
                         STATION_NAME,dist_error, dist.m) %>% 
  nest() %>% ungroup()

# loop through each objectid and station number pair to calculate initial sill elevations
combo.all = NULL
for (y in 1:nrow(nest.sill)){
  df = nest.sill$data[[y]]
  obj_id = nest.sill$OBJECTID[[y]]
  stat_id = nest.sill$ID[[y]]
  stat_nam = nest.sill$STATION_NAME[[y]]
  dist_error = nest.sill$dist_error[[y]]
  dist_m = nest.sill$dist.m[[y]]
  n.obs = nrow(df)
  obs.count = df %>% group_by(.pred_class) %>% count() %>% ungroup() %>% 
    mutate(all.obs = n.obs,
           pct = n/n.obs)
  
   # cond1 = isTRUE(obs.count$pct[obs.count$.pred_class==0]>=0.95) &
   #   nrow(df[df$.pred_class==0,])>=5
 
  
  if(isTRUE(obs.count$pct[obs.count$.pred_class==0]>=0.95) &
     nrow(df[df$.pred_class==0,])>=5  ){
    class = "always 0"
    combo = cbind.data.frame(obj_id, class, stat_id, stat_nam, dist_m, dist_error, 
                           num.0=NA, mean.0=NA, sd.0=NA, min.0=NA, max.0=NA, 
                           num.2=NA, mean.2 = NA, sd.2 =NA, min.2 = NA, max.2 = NA, 
                           pval=NA)
    combo.all = rbind.data.frame(combo.all, combo)
    next
  }
  if(isTRUE(obs.count$pct[obs.count$.pred_class==2]>=0.95)& nrow(df[df$.pred_class==2,])>=5){
    class = "always 2"
    combo = cbind.data.frame(obj_id, class, stat_id, stat_nam, dist_m, dist_error, 
                           num.0=NA, mean.0=NA, sd.0=NA, min.0=NA, max.0=NA, 
                           num.2=NA, mean.2 = NA, sd.2 =NA, min.2 = NA, max.2 = NA, 
                           pval=NA)
    combo.all = rbind.data.frame(combo.all, combo)
    next
  }
  if(isTRUE(nrow(df[df$.pred_class==2,])<5) | isTRUE(nrow(df[df$.pred_class==0,])<5)){
    next
  }
  ttest = t.test(df[df$.pred_class == 0,]$Value.0, df[df$.pred_class == 2,]$Value.0 )
  pval = ttest$p.value
  if(pval>0.05){
    class = "no discharge relationship"
    combo = cbind.data.frame(obj_id, class, stat_id, stat_nam, dist_m, dist_error, 
                           num.0=NA, mean.0=NA, sd.0=NA, min.0=NA, max.0=NA, 
                           num.2=NA, mean.2 = NA, sd.2 =NA, min.2 = NA, max.2 = NA, 
                           pval=pval)
    combo.all = rbind.data.frame(combo.all, combo)
    
    next}
  num.0 = df[df$.pred_class == 0,] %>% nrow()
  num.2 = df[df$.pred_class == 2,] %>% nrow()
  mean.0 = mean(df[df$.pred_class==0,]$Value.0)
  sd.0 = sd(df[df$.pred_class==0,]$Value.0)
  min.0 = min(df[df$.pred_class==0,]$Value.0)
  max.0 = max(df[df$.pred_class==0,]$Value.0)
  mean.2 = mean(df[df$.pred_class==2,]$Value.0)
  sd.2 = sd(df[df$.pred_class==2,]$Value.0)
  min.2 = min(df[df$.pred_class==2,]$Value.0)
  max.2 = max(df[df$.pred_class==2,]$Value.0)
  class = "discharge dependant"
  combo = cbind.data.frame(obj_id, class, stat_id, stat_nam, dist_m, dist_error, 
                           num.0, mean.0, sd.0, min.0, max.0, 
                           num.2, mean.2, sd.2, min.2, max.2, 
                           pval)
  combo.all = rbind.data.frame(combo.all, combo)
}

setwd(int.wd)
write_feather(combo.all, paste0("raw_sillElevation_material.feather"))
combo.all = read_feather("raw_sillElevation_material.feather")

# calculate sill elevation ranges and the mid point of that range
combo.sill = combo.all %>% 
  mutate(xmin = case_when(
                       max.0>min.2 ~ min.2-as.numeric(dist_error), 
                       max.0<min.2 ~ max.0-as.numeric(dist_error)
                     ),
         xmax = case_when(
           max.0>min.2 ~ max.0+as.numeric(dist_error),
           max.0<min.2 ~ min.2 + as.numeric(dist_error)
                    ),
         mid.sill = (xmin+xmax)/2) %>% 
  as_tibble() %>% 
  select(obj_id, class, stat_id, stat_nam, dist_m, dist_error, xmin, xmax, mid.sill, pval, num.0, num.2) %>% 
  as_tibble()

# calculate ranges for sills. If two stations produced two different sill groups for a lake, take a peak at what is going on. Keep the class from the station that is closest to the lake
diffclass.diffstat = combo.sill %>%
  group_by(obj_id, class) %>% count() %>% ungroup() %>% 
  group_by(obj_id) %>% count() %>% ungroup()%>% 
  rename(numclasses = n)
keep.obs = combo.sill %>% 
  left_join(diffclass.diffstat, by="obj_id") %>% as_tibble() %>% 
  filter(numclasses>1) %>% 
  mutate(classFactor = factor(class, levels = c("discharge dependant", "always 0", "always 2", "no discharge relationship"))) %>% 
  group_by(obj_id,classFactor, class) %>% count() %>% ungroup() %>% 
  arrange(obj_id, classFactor) %>% 
  group_by(obj_id) %>% filter(row_number()==1) %>% 
  select(obj_id, class) %>% 
  mutate(keep.type = "keep")
combo.prep = combo.sill %>% 
  left_join(diffclass.diffstat, by="obj_id") %>% 
  left_join(keep.obs, by=c("obj_id", "class")) %>% as_tibble() %>% 
  filter(numclasses == 1 | (numclasses >1 & !is.na(keep.type)))
  


# for the lakes that are in 50km of multiple stations that all are able to calculate sill elevation ranges, compare the sill elevation ranges to see if they overlap
compare.sills = combo.prep%>% as_tibble() %>% 
  filter(!is.na(mid.sill)) %>%
  select(obj_id, stat_id, dist_m, mid.sill, xmin, xmax) %>% 
  group_by(obj_id) %>% nest()
return_all = NULL
for (b in 1:nrow(compare.sills)){
  sill.df = compare.sills$data[[b]] %>% arrange(dist_m) %>% as.data.table()
  obj_id = compare.sills$obj_id[[b]]
  len = nrow(sill.df)
  results = cbind.data.frame(obj_id, len, sill.df[, .(max(xmin), min(xmax))])
  return_all = rbind.data.frame(return_all, results)
}
sill.compare = combo.prep %>% 
  left_join(return_all, by="obj_id") %>% 
  mutate(diff = V2-V1,
         fin.min = case_when(
           !is.na(V1) & diff>0 ~ V1,
           !is.na(V1) & diff<0 ~ -999
         ),
         fin.max = case_when(
           !is.na(V2) & diff>0 ~ V2,
           !is.na(V2) & diff<0  ~ -999
         )) %>% 
  select(obj_id, class,fin.min, fin.max, stat_id) %>%arrange(obj_id) %>% 
  group_by(obj_id) %>% 
  mutate(rnm = row_number()) %>% 
  spread(rnm, stat_id) %>% 
  rename(STATION_1 = `1`, STATION_2 = `2`, STATION_3 = `3`, STATION_4 = `4`) %>% 
  mutate(fin.range = ifelse(fin.max == -999 | 
                              fin.min ==-999, NA, 
                            fin.max-fin.min),
       fin.sill = ifelse(fin.max == -999 | 
                              fin.min ==-999, NA, 
                            (fin.max+fin.min)/2)) 


setwd(int.wd)
write_feather(sill.compare,paste0(todayDate, "sillElevation_c.feather"))
#sill.compare = read_feather(paste0(todayDate, "sillElevation_c.feather"))

```

# Plot sill elevation figures
## Figure 7
```{r}
setwd(int.wd)
sill.compare = read_feather(paste0(todayDate, "sillElevation_c.feather"))

# Join with lat/lon data
sill.sf = sill.compare %>% rename(OBJECTID=obj_id) %>% left_join(lakes.sf, by="OBJECTID") %>% 
  st_as_sf() %>% st_transform(crs.plot)

station.locations.sill = level.locations %>% 
  filter(ID != "10MC002" &  ID != "10LC014" &
           ID != "10LC021" & ID != "10MC008")

# Plot Figure 7
p1.all=ggplot()+
  geom_sf(data=sill.sf %>% filter(!is.na(fin.sill)),
          aes(fill=fin.sill), color=NA)+
  scale_fill_viridis_c(option="inferno", limits = c(-0.25, 4))+
  theme_void()+
  annotation_scale()+
  geom_sf(data=mack.basin.large, color="grey70", size=0.5)+
 # geom_sf(data=station.locations.sill, color="black",size=2, shape=17)+
  theme(
        panel.background = element_rect(fill=NA, color=NA),
        legend.position="bottom",
        legend.text = element_text(size=12),
        legend.background = element_blank(),
        legend.title = element_text(size=12, face="bold"))+
  labs(fill="median\nfunctional\nconnectivity\nelevation\nthreshold (m)")+
   guides(fill = guide_colorbar(barwidth = 8, barheight = 0.5))+ggtitle("a.")
p1.all

p2.all = ggplot()+
  geom_sf(data=sill.sf %>% filter(!is.na(fin.sill)),
          aes(fill=fin.range/2), color=NA)+
  scale_fill_viridis_c(option="viridis", limits=c(0,3))+
  geom_sf(data=mack.basin.large, color="grey70", size=0.5)+
  #geom_sf(data=station.locations.sill, color="black",size=2, shape=17)+
  theme_void()+
  theme(
    panel.background = element_rect(fill=NA, color=NA), 
    legend.background = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size=12),
    legend.title = element_text(size=12, face="bold"))+
  labs(fill="uncertainty in\nfunctional\nconnectivity\nelevation threshold\n(m)")+
   guides(fill = guide_colorbar(barwidth = 8, barheight = 0.5))+ggtitle("b.")
p2.all

p3.all=ggplot()+
  geom_sf(data=sill.sf %>% filter(is.na(fin.sill)) %>% 
            mutate(class = ifelse(class=="discharge dependant", 
                                  "high uncertainty,\ncan't calculate\nelevation threshold", 
                                  class)) %>% 
            mutate(class = ifelse(class=="no discharge relationship", 
                                  "no significant\nwater level\nrelationship", class)),
          aes(fill=class), color=NA)+
  #scale_fill_viridis_c(option="viridis", limits = c(0,1))+
  geom_sf(data=mack.basin.large, color="grey70", size=0.5)+
 # geom_sf(data=station.locations.sill, color="black",size=2, shape=17)+
  theme_void()+
  theme(
   
    legend.background = element_blank(),
    
    panel.background = element_rect(fill=NA, color=NA), 
    legend.position = "bottom",
    legend.text = element_text(size=12),
    legend.title = element_blank())+
   guides(fill=guide_legend(nrow=4,byrow=TRUE))+ggtitle("c.")


p1.all+p2.all+p3.all
setwd(images.wd)
ggsave(paste0("20231018","_sills_ALL.png"), width=10.5, height=8.5, units = "in")
ggsave(paste0("20231018","_sills_ALL.pdf"), width=10.5, height=8.5, units = "in")

# table with a summary of lakes
sill.compare %>% lazy_dt() %>% 
    mutate(final.class = case_when(
    class=="always 0" ~ "always 0",
    class=="always 2" ~ "always 2",
    class=="discharge dependant" & (fin.range<=1) ~ 
      "discharge dependant (<1m uncert)",
    class=="discharge dependant" & (fin.range>1 & fin.range<=1.5) ~ 
      "water level dependant (1-1.5m uncert)",
    class=="discharge dependant" & (fin.range>1.5 & fin.range <=2) ~ 
      "water level dependant (1.5-2m uncert)",
    class=="discharge dependant" & (fin.range>2) ~ 
      "water level dependant (>2m uncert)",
    class=="discharge dependant" & (is.na(fin.range)) ~ 
      "water level dependant (unable to calculate sill)",
    class=="no discharge relationship" ~ "no discharge relationship"
  )) %>% group_by(final.class) %>% count()
```


# Comparison of our sill elevations to Marsh & Hey 1988
## Figure 8 & Figure 9
```{r}
# Import the sill elevations calculated above
setwd(int.wd)
sill.compare = read_feather(paste0(todayDate, "sillElevation_c.feather"))

# import the validation and join it to the same object ids as used above
setwd(val.wd)
val.raw = read_sf(val.filename) %>% st_transform(crs.plot)
lakes.prj = lakes.sf %>% st_transform(crs.plot)


combo.df = val.raw %>% st_join(lakes.prj %>% select(geometry, OBJECTID)) %>% 
  filter(!is.na(OBJECTID)) %>% 
  left_join(sill.compare %>% select(obj_id, fin.range, fin.sill, class) %>% 
              rename(OBJECTID = obj_id), by="OBJECTID") %>% mutate(errorbar = fin.range/2)

# Plot Figure 7
plot1 = combo.df %>%  ggplot(aes(x=fin.sill, y=SmmrSill))+
  geom_errorbarh(aes(xmin = fin.sill-errorbar, 
                     xmax = fin.sill+errorbar, y=SmmrSill, height=0.1), 
                 alpha=0.5, color="grey50")+
  geom_errorbar(aes(x=fin.sill, ymin=SmmrSill-0.5, 
                    ymax =SmmrSill +0.5, width=0.1), alpha=0.5, color="grey50")+
  theme_bw()+
  geom_abline(aes(slope=1, intercept=0), lty=2, color="grey60")+
  geom_point(color="red")+
  geom_point(data=combo.df %>% filter(errorbar<=0.5), aes(x=fin.sill, y=SmmrSill), color="black")+
 # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on")+
  xlab("functional connectivity elevation threshold (m)")+
  ylab("Marsh & Hey (1988)\nSummer Sill Elevation (m)")+xlim(0,5)+ylim(0,5)+
  geom_text(aes(x=4.2, y=4.4, label="1:1"), angle=45, color="grey60")
plot1
setwd(images.wd)
ggsave("MarshHeyAll_versusPlot.png", width=4, height=4, units = "in")


# calculate whether or not the sills overlap for always 0 and alway 2 functional connectivity lakes
combo.filt = combo.df %>% na.omit() %>% 
  mutate(sill_lowerBound = SmmrSill-0.5,
         sill_upperBound = SmmrSill + 0.5,
         thresh_lowerBound = fin.sill - errorbar,
         thresh_upperBound = fin.sill + errorbar)


combo.prep = combo.df %>% mutate(group_name = case_when(
   class=="always 0" ~ "always low functional connectivity (class 0)",
   class=="always 2" ~ "always high functional connectivity (class 2)",
   class=="no discharge relationship" ~ "no discharge relationship"
 ))

combo.prep$group_name = 
  factor(combo.prep$group_name,levels = 
           c("always low functional connectivity (class 0)",
             "always high functional connectivity (class 2)",
             "no discharge relationship"))

# Figure 9
plot2 = combo.prep %>%  
  filter(is.na(fin.sill) & class!="no discharge relationship") %>%
  ggplot(aes(x=str_wrap(group_name, 12), 
             y=SmmrSill))+geom_boxplot()+
  theme_bw()+xlab("")+theme(legend.position="none")+
  ylab("Marsh & Hey (1988)\nSummer Sill Elevation (m)")+
  labs(fill="Functional\nConnectivity Class")
plot2
setwd(images.wd)
ggsave("MarshHeyAll_noSillElev.png", width=3, height=3, units = "in")

# Plot summary table
combo.df %>% as_tibble() %>% select(-geometry) %>% mutate(class=case_when(
  is.na(fin.sill) ~ class,
  !is.na(fin.sill) & fin.range <=1 ~ "low error (<= +/- 0.5m) functional sill",
   !is.na(fin.sill) & fin.range >1 ~ "high error (> +/- 0.5m) functional sill"
)) %>% group_by(class) %>% count()

```


# Calculate connectivity durations for Table 5
```{r}
# import sill elevation data
setwd(int.wd)
sill.compare = read_feather(paste0(todayDate, "sillElevation_c.feather")) %>% 
  filter(!is.na(fin.sill)& fin.range < 1) %>% # keep only lakes that have a sill elevation with a range less than 1m
  rename(OBJECTID=obj_id)

# Select only water level data from years where we have mostly complete records (at least 113/153 days) at at least 5/6 stations
keep.years = level_data %>% 
  filter(ID != "10MC002" &  ID != "10LC014" &
           ID != "10LC021" & ID != "10MC008") %>% 
  filter(!is.na(Value)) %>% filter(month>=5) %>% 
  group_by(ID, year) %>% count() %>% ungroup() %>% 
  filter(n>113) %>% #153 is max days, only allow up to 20 missing days
  group_by(year) %>% count() %>% filter(n>=5) %>%  ungroup()

# select years for each station
keep.station.years = level_data %>% 
  filter(ID != "10MC002" &  ID != "10LC014" &
          ID != "10LC021" & ID != "10MC008") %>% 
  filter(!is.na(Value)) %>% filter(month>=5) %>% 
  group_by(ID, year) %>% count() %>% ungroup() %>% 
  filter(n>113 & year %in% keep.years$year)

# for those selected years, calculate the average water level at each station
mean.level = level_data %>% 
  filter(ID != "10MC002" &  ID != "10LC014" &
           ID != "10LC021" & ID != "10MC008") %>% 
  left_join(keep.station.years, by=c("ID", "year")) %>% 
  filter(month>=5) %>% 
  filter(!is.na(n)) %>% 
  group_by(ID,doy) %>% 
  summarise(mean.level = mean(Value, na.rm=T)) %>% as.data.table()

N <- nrow(sill.compare)
above.min <- vector("list", N)
above.mid = vector("list", N)
above.max = vector("list", N)

for (z in 1:nrow(sill.compare)){
  df = sill.compare[z,]
  stations = df %>% select(starts_with("STATION"))
  stations = as.data.frame(t(stations))$V1 
  fin.min = df$fin.min[1]
  fin.max = df$fin.max[1]
  fin.sill = df$fin.sill[1]
  
  level_df = mean.level[ID %in% stations]
  level_df = dcast(level_df, doy~ID, value.var="mean.level")
  level_df = na.omit(level_df)
   
  relevant.level.mean = level_df[, .(Mean = rowMeans(.SD)), by = doy]
   
  above.min.num = relevant.level.mean[Mean>=fin.min, .N]
  above.mid.num = relevant.level.mean[Mean>=fin.sill, .N]
  above.max.num = relevant.level.mean[Mean>=fin.max, .N]
  
  above.min[[z]] = above.min.num
  above.mid[[z]] = above.mid.num
  above.max[[z]] = above.max.num
}
above.min = above.min %>% unlist()
above.mid = above.mid %>% unlist()
above.max = above.max %>% unlist()
OBJECTID = sill.compare$OBJECTID
range = sill.compare$fin.range
sill.days = cbind.data.frame(OBJECTID, range, above.min, 
                             above.mid, above.max) %>% as_tibble()

setwd(int.wd)
write_feather(sill.days, paste0(todayDate, "avgConnectionTime.feather"))
#sill.days=read_feather(paste0(todayDate, "avgConnectionTime.feather"))

# Plot values used in Table 5
setwd(int.wd)
read_feather(paste0(todayDate, "avgConnectionTime.feather")) %>% 
  mutate(group.mid = case_when(
    above.mid <=14 ~ "0-14 days",
    above.mid>=15 & above.mid <=60 ~ "15-60 days",
    above.mid>=61 ~ "above 61 days"
  )) %>% group_by(group.mid) %>% count()

read_feather(paste0(todayDate, "avgConnectionTime.feather")) %>% 
  mutate(group.max = case_when(
    above.max <=14 ~ "0-14 days",
    above.max>=15 & above.max <=60 ~ "15-60 days",
    above.max>=61 ~ "above 61 days"
  )) %>% group_by(group.max) %>% count()


read_feather(paste0(todayDate, "avgConnectionTime.feather")) %>% 
  mutate(group.min = case_when(
    above.min <=14 ~ "0-14 days",
    above.min>=15 & above.min <=60 ~ "15-60 days",
    above.min>=61 ~ "above 61 days"
  )) %>% group_by(group.min) %>% count()

```


# Plot gif of results w/ discharge
# not in paper
```{r}
setwd("C:/Users/whyana/OneDrive/DocumentsLaptop/001_GraduateSchool/Research/Connectivity/Mackenzie/images/GIF_20230512")
prep.gif.data = filt.obs %>%  # all lakes during 4 weeks after freshet
  left_join(lakes.sf, by="OBJECTID") %>% 
  st_as_sf() %>% st_transform(crs = crs.plot) %>% 
  group_by(year) %>% nest() %>% ungroup()

dis.location = hy_stations(station_number = "10LC014") %>% 
  st_as_sf(coords=c("LONGITUDE", "LATITUDE")) %>% 
  st_set_crs(4326) %>% st_transform(crs.plot)
dis.filt = complete.flows %>% filter(STATION_NUMBER=="10LC014")

study.area.gif=cbind.data.frame(lon=c(-137.3, -137.3, -133.2, -133.2), 
                 lat=c(67.25, 69.64, 69.64, 67.25)) %>% 
  st_as_sf(coords=c("lon", "lat")) %>% st_set_crs(4326) %>% st_bbox() %>% st_as_sfc() %>% 
  st_transform(crs = crs.plot)

### Loop through each year
for (z in 1: length(prep.gif.data$year)){
  dat = prep.gif.data$data[[z]] 
  year.main = prep.gif.data$year[[z]]
  
  #### Create plot 1 (map of June connectivity)
  scale = scale_fill_gradientn(colours = c("#88ccee","#44aa99","#117733"), limits=c(0,2))
  p1 = ggplot(data=dat)+
   geom_sf(aes(fill=mean.con), color=NA)+
    theme_bw()+scale+
    annotation_scale(text_cex = 1.2)+
    geom_sf(data=mack.basin.large, color="grey65")+
    geom_sf(data=study.area.gif , color=NA, fill=NA)+
    #geom_sf(data=dis.location,color="black", size=5)+
    scale_colour_manual(guide="none", values=c("#000000", "#ABA9A9"))+
    ggtitle(year.main)+
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          plot.title=element_text(size=18, face="bold", hjust=0.5),
         # legend.title = element_text(size=16, face="bold"),
          legend.text=element_text(size=16, face="bold"),
          legend.position="bottom", legend.direction="horizontal",
          legend.key.size=unit(1, "cm"),
          axis.ticks = element_blank(),
          legend.title=element_blank(),
         legend.box.spacing = unit(0, "pt"),
         legend.margin=margin(0,0,0,0))+labs(fill="Class")+
    guides(fill=guide_legend(label.position="top",label.vjust = -8, title.vjust = 0.2))
  

  #### Save the gif to your file
  ggsave(plot=p1,filename= paste0("year", year.main, ".png") ,width=14, height=9.5, units = "in") 
}  


### List all the files in the gif and combine them into a stacked image

files=list.files(pattern="*.png")
images <- map(files, image_read)
images <- image_join(images)
### Animate the stacked image
gif = image_animate(images, fps = 1, dispose = "previous")

## save as a gif
setwd(images.wd)

image_write(gif, paste0("gif_", "20230512", ".gif"))


```

# Compare sediment to water level -- what is the relationship?
## not in paper, but useful for presentations or reviewer's comments
```{r}
# impart Mackenzie Sediment / Water Quality data from Arctic Red River
setwd(sed.wd)

wq.df = read.csv(sed.file) %>% 
  as_tibble() %>% select(Date, Discharge, DOC, TSS) %>% 
  mutate(Date = as_date(Date))

# Filter water level data just to Arctic Red River
rr.water.level = level_data %>% filter(ID == "10LC014") %>% 
  mutate(Date = as_date(Date))

# join discharge and water quality data
wl.wq = wq.df %>% left_join(rr.water.level, by="Date") %>% 
  filter(!is.na(TSS) & !is.na(Value))

ggplot(wl.wq)+
  geom_point(aes(x=Value, y=TSS, color=as.factor(month)))+
  xlab("Water Level (m) at Arctic Red River")+ylab("TSS (mg/L) at Arctic Red River")+theme_classic()+labs(color="Month")+
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14, face="bold"),
        legend.text = element_text(size=14),
        legend.title = element_text(size=14, face="bold"),
        panel.background = element_rect(fill='transparent'),
        #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA),
    #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), 
    #transparent legend bg
    legend.box.background = element_rect(fill='transparent', color=NA)) 
#transparent legend panel)
setwd(images.wd)
ggsave("20230508_TSS_vs_waterLevel.png", bg="transparent",
       width = 6, height = 4, units="in")
# TLDR: Above the delta, water level and TSS are highly correlated. 
```

